Lightness constancy under fixed geometry

### 2017-10-14 Draft1/draft1.tex
edit abstract

\title{Lightness constancy under fixed geometry}
\abstract{Color constancy is the ability of human visual system to stably perceive the color of an object under variations in the light reflected from the object. This is a very significant invariance detection ability with important behavioral implications. The light captured by our visual system depends not just on the object of interest, but also the lighting and other objects in the surround. In order to corectly identify the object color, the variations resulting from the surround related factors must be discarded. To understand the computational principles that can lead to color constancy, we study constancy of the lightness component of the 3D color space. Lightness is the attribute of the color space that determines the brightness of the color. We study the effects of spectral variations, specifically the variations in the object reflectance spectra and the light source illumination spectra, on the perceived lightness of an object. We have employed a supervised learning approach to study lightness constancy. We have developed a computational software that generates naturalistic hyper-spectral images of 3D scenes. The software allows precise control of the geometrical and spectral properties of the objects and lights in the scene. It takes into account the statistical structure of the spectral variations in natural scenes. It also produces the response of the retinal cones to the hyperspectral images through an accurate model of human visual system. We label the iamges generated by the software through the standard lightness (lighness under standard daylight spectrum D65) of a specific target object in the scene. We study the corresponsding dataset of labeled cone responses using supervised computational learning methods. Our analysis shows that under simple spectral variations, like the variation only in the spectrum of the target object, or of the background object, or of the light source, the standard lightness can be recovered by simple trasnformations of the reflected light. When all spectra vary simultaneously, while it is not easy to recover the lightess through simple transformations, a decoding cheme that compares the light from the target and the surround and properly adds the response of the L,M and S cones can recover the lightness within about 15\% RMSE. }

@@

\section{Introduction}
The color of an object has important behavioral implications. This important property of an object is perceived from the light reflected off the object surface. The reflected light depends not just on the properties intrinsic to the object, such as the object's surface reflectance, its texture, material properties and geometry (shape, size, orientation), but also on extrinsic factors that depend on the object's surround (Fig. \ref{fig:introSchematic}). The surround dependent extrinsic factors that affect the reflected light are: the surface reflectance, texture and material properties of the objects in the surround, the illumination spectrum of the sources of light illuminating the object, the geometry, relative position and pose of the objects and light sources in the surround, and the position of the observer. As these extrinsic factors vary considerably from scene to scene, the challenge faced by the visual system is to recover a stable perceptual representation of the object's surface reflectance, ignoring the  variations in the reflected light resulting from the extrinsic factors. The ability of the visual system to extract such a stable representation, its perceived color, is called color constancy. We are interested in identifying the computations that lead to such constancy.

In this work, we study a simpler analogue of color constancy called lightness constancy. Lightness is the attribute that determines the brightness of a color. A color with a low lightness is nearly black, while one with a high lightness is the pure, saturated color. For achromatic objects, it is the property that changes from black to gray to white. Here, we only consider the lightness component of the  3D human perceptual color space and study the effects of variations in the reflected light on the perceived lightness.

@@

The variations in the reflected light can result from either geometrical factors (like shape, size, relative position of objects, etc.) or spectral factors (such as the illumination spectra, surface reflectance, diffusivity, etc.). In this work, we focus on the spectral factors. Specifically, we study the effects of variations in the surface reflectance spectra and the illumination spectra on lightness perception  (Fig.\ref{fig:introExampleFigure}). We are interested in identifying the computations that can lead to stable lightness perception.

@@

\section{Results}
\subsection{Case 1}

@@

\subsection{Lightness estimates and model performance}
Fig. ZZ gives the AMA lightness estimates for the three cases discussed above. We show the target object lightness estimates for there different methods: linear regression on center pixel, SVD regression and AMA. As is expected, when only the target object reflectance varies, it is easy to estimate the target object lightness. For case 2, where the light source spectra are allowed to vary, although the light captured by the cones vary from image to image, the contrast between the target and the background remains fixed and can be used to estimate the target object lightness. Thus, the SVD-LDA and AMA still provide good estimates of the target object lightness. In case 3, the most challenging case where all the spectra are allowed to change, AMA still does a reasonable job of estimating the target object lightness. AMA mean estimates lie close to the diagonal x=y line.

>>>>

\title{Lightness constancy under fixed geometry}
\abstract{Human visual system has the ability to stably perceive the color of an object under variations in the reflected light due to surround factors. This is a behaviorally significant invariance detection ability whose biological mechanism remains unknown. To understand the computational principles that can lead to such invariant perception, here we study the perception of object lightness under spectral variations. Lightness is the attribute of the 3D color space that determines the brightness of a color. Specifically, we study the effects of variations in reflectance and illumination spectra of the objects and lights sources in a scene on the perceived lightness of an object. We have developed a software to generate naturalistic hyper-spectral images of 3D scenes with precise control over the geometrical and spectral properties of the objects and lights in the scene. The software also produces the response of the early visual system to these hyper-spectral images. We label the images with the lightness of a specific object in the scene and use supervised learning methods to identify the computations that lead to stable lightness perception. Our analysis shows that for simple spectral variations, like the ones that allow variations only in the spectrum of the target object, or of the background object, or of the light source, the standard lightness can be recovered by through transformations of the reflected light. When all spectra are allowed to vary simultaneously, while it is not easy to recover the lightness through simple transformations, a decoding sheme that compares the light from the target and the surround, and properly adds the response of the L,M and S cones can recover the lightness within about 15\% RMSE.}

@@

\section{Introduction}
The color of an object has important behavioral implications. This important property of an object is perceived from the light reflected off the object surface. The reflected light depends not just on the properties intrinsic to the object, such as the object's surface reflectance, its texture, material properties and geometry (shape, size, orientation), but also on extrinsic factors that depend on the object's surround (Fig. \ref{fig:introSchematic}). These extrinsic factors are: the surface reflectance, texture and material properties of the objects in the surround, the illumination spectrum of the sources of light illuminating the object, the geometry, relative position and pose of the objects and light sources in the surround, and the position of the observer. As these extrinsic factors vary considerably from scene to scene, the challenge faced by the visual system is to recover a stable perceptual representation of the object's surface reflectance, invariant of the extrinsic factors. The ability of the visual system to extract such a stable representation, its perceived color, is called color constancy. We are interested in identifying the computations that lead to such constancy.

In this work, we identify the computations that lead to stable lightness perception. Lightness is the attribute that determines the brightness of a color. A color with a low lightness is nearly black, while one with a high lightness is the pure, saturated color. For achromatic objects, lightness is the property that changes from black to gray to white. We study the effects of variations in the reflected light on the perceived lightness.

@@

The variations in the reflected light can result from either geometrical factors (like shape, size, relative position of objects, etc.) or spectral factors (such as the illumination spectra, surface reflectance, diffusivity, etc.). Here, we focus on the spectral factors. Specifically, we study the effects of variations in the surface reflectance spectra and the illumination spectra on lightness perception  (Fig.\ref{fig:introExampleFigure}). We are interested in identifying the computations that can lead to stable lightness perception.

@@

\section{Results}
\subsection{Case 1}
Fig.~\ref{fig:case9AllResults} shows the receptive fields and the the AMA lightness estimates for the three cases discussed above. We show the target object lightness estimates for there different methods: linear regression on center pixel, SVD regression and AMA. As is expected, when only the target object reflectance varies, it is easy to estimate the target object lightness. For case 2, where the light source spectra are allowed to vary, although the light captured by the cones vary from image to image, the contrast between the target and the background remains fixed and can be used to estimate the target object lightness. Thus, the SVD-LDA and AMA still provide good estimates of the target object lightness. In case 3, the most challenging case where all the spectra are allowed to change, AMA still does a reasonable job of estimating the target object lightness. AMA mean estimates lie close to the diagonal x=y line.

@@
